---
title: "Mini Project"
author: "S/18/824"
date: "2024-04-03"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the necessary packages
```{r}
library(tidyverse)
library(psych)
library(GPArotation)
library(lavaan)
```

Import the data set
This data set have 16 columns.I consider only numerical columns for the analysis 
```{r}
dry_bean <- read_csv("../data/Dry_Bean_Dataset.csv",col_types = "idddddiddddddddd")
head(dry_bean)
```
```{r}
glimpse(dry_bean)
```
##This dataset have 16 columns and 13611 rows.
```{r}
str(dry_bean)
```


Check for missing values
```{r}
sum(is.na(dry_bean))
```

Describe the data set
```{r}
describe(dry_bean)
```

Standardized the data set

```{r}
std_dry_bean <- scale(dry_bean)
std_dry_bean
```

Get the correlation matrix of standardized data
```{r}
dry_bean_cor <- cor(std_dry_bean)
dry_bean_cor
```

Find the eigen values and eigen vectors of above correlation matrix
```{r}
dry_bean_eigen <- eigen(dry_bean_cor)

#Eigen values
dry_bean_eigen$values

#Eigen vectors
dry_bean_eigen$vectors
```

##According to the eigen values we select number of factor as 3.
##Because first 3 eigen values are greater than 1.

Proportion of variance explained
```{r}
prop_var <- dry_bean_eigen$values/sum(dry_bean_eigen$values)
prop_var
```
Draw the scree plot
```{r}
scree(dry_bean)
```

##From the above scree plot after dimension 4 graph gradually bendsoff, so we select number of factor as 3 
##That is adequately retained the dataset
Parallel analysis Scree Plots
```{r}
fa.parallel(dry_bean,fm="pa",fa="fa")
```
##From the scree plot we can see that number of factors to be 3.

a.) Factor Analysis from 'Principal Component Method'
```{r}
dry_bean_pc <- fa(dry_bean_cor,nfactors = 3,rotate = FALSE,fm = "pa")
dry_bean_pc
```
Get un rotated loadings from PC method
```{r}
print(dry_bean_pc$loadings,cutoff = 0.3)
```
##From the above results we can see that some variables are cross loading with more principal components.
##So, to identify the factors clearly we need to rotate it inorder to get the clear interpretation.
##About 91 % of total variation is explained by the 3 factors.

Get unrotated PC method communalities 
```{r}
dry_bean_pc$communalities
```

Rotate the PC method factor loadings using 'Varimax' method
```{r}
dry_bean_pc_rotate <- fa(dry_bean_cor,nfactors = 3,rotate = "promax",fm = 'pa')
dry_bean_pc_rotate
```
```{r}
print(dry_bean_pc_rotate$loadings,cutoff = 0.5,sort = TRUE)
```
```{r}
model <- '
DM =~Area+Perimeter+MajorAxisLength+MinorAxisLength+ConvexArea+EquivDiameter+ShapeFactor1
RM =~AspectRation+Eccentricity+roundness+Compactness++ShapeFactor2+ShapeFactor3
PM =~Solidity'

fit <- cfa(model,data = dry_bean)


summary(fit,fit.measures = TRUE,standardized = TRUE)
```
Since variances of some variables are very high, and dataset doesnot converge.So, We can use some other techniques. 

Get rotated communalities in PC method
```{r}
dry_bean_pc_rotate$communalities
```

b.) Factor Analysis from 'Maximum Likelihood Method'
```{r}
dry_bean_ml <- fa(dry_bean_cor,nfactors = 3,rotate = FALSE, fm = 'ml')
dry_bean_ml
```
Get unrotated ML loadings 
```{r}
print(dry_bean_ml$loadings)
```
##Here, we can't clearly understand the factors because some variables are crossed with more than 2 factors.So we need to rotate it.

Get unrotated Communalities in ML method
```{r}
dry_bean_ml$communalities
```
Rotate the ML method factor loadings using 'Varimax' method
```{r}
dry_bean_ml_rotate <- fa(dry_bean_cor,nfactors = 3,rotate = "varimax",fm = 'ml')
dry_bean_ml_rotate
```

Get rotated ML loadings
```{r}
print(dry_bean_ml_rotate$loadings,cutoff = 0.5)
```
##From this about 81.9 % of total variation is explained by these factors.

```{r}
dry_bean_ml_rotate$communalities
```

